\textsf{Are large language models superhuman chemists?}

### 2024-10-18 src/tex/appendix.tex
more cleanup

\subsection{Benchmark corpus}
To ensure maximal interoperability with existing benchmarks or tools, we curated the data in an extended form of the widely used BigBench format.\autocite{srivastava2022beyond}
This also implies that future baselines can be built on top of our infrastructure if saved in the same format.

\Cref{fig:flesch_kincaid_reading_ease} shows the distribution of the Flesch-Kincaid reading ease scores of the questions.
We see that the questions are generally complex to read.

@@

        \caption{Examples for each of the topics considered in the evaluation of the \chembench corpus.}
    \begin{tabularx}{\textwidth}{p{2.7cm}p{2cm}X}

@@

@@

\begin{table*}[!t]
    \centering
        \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.}}
    \begin{tabularx}{\textwidth}{p{2.7cm}p{2cm}X}
        \toprule
        Required Skills & Number of Questions & Example Question \\
        \midrule
        Knowledge & 770 & Which of the following salts is an acidic salt? \\
        & & A. \ce{NH4Cl} \\
        & & B. \ce{Na2CO3} \\
        & & C. \ce{NaH2PO4} \\
        & & D. \ce{Zn(OH)Cl} \\
        \midrule
        Reasoning & 825 & How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
        & & A. By measuring atomic force microscopy \\
        & & B. By measuring fourier transform infrared spectroscopy \\
        & & C. By measuring photoluminescence spectrum \\
        & & D. By measuring the absorption spectroscopy \\
        \midrule
        Intuition & 1001 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality). Which of the following two candidates would you prefer for further development? \\
        & & A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
        & & B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC(=O)c2ccco2)CC1[END\_SMILES] \\
        \midrule
        Calculation & 91 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution. The volume of \ce{KOH} solution required to reach the endpoint was \SI{6.81}{cm^3}. What is the average number of monomer units in each polymer chain of this sample? \\
        \bottomrule
    \end{tabularx}
    \label{tab:chembench_corpus_cognitive}
\end{table*}

% \subsection{Parsing verification}\label{sec:manually-verified-parsing}
% For validating the parsing workflow, we randomly sampled four questions per topic and manually verified that the completions of the model were parsed correctly.

@@

\begin{table}
    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
    \resizebox{\textwidth}{!}{
    \variable{output/performance_table.tex}
    }
\end{table}

\begin{table}
    \caption{\textbf{Performance of the models on the \enquote{tiny} subset of the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
    \resizebox{\textwidth}{!}{
    \variable{output/performance_table_human_subset.tex}
    }
\end{table}

@@

\subsection{Implementation}

\begin{figure}
    \centering
    \includegraphics[height=\textheight]{figures/process.pdf}
    \caption{\textbf{Overview of the benchmarking pipeline implemented in \chembench.} The process begins with JSON files containing task data, which are used to create \texttt{Task} objects and initialize the \texttt{ChemBenchmark}. \texttt{Prompter} objects are then created to handle different types of prompts for instruction-tuned and completion models.
    The \texttt{Task} Iterator differentiates between \gls{mcq} and other question types. For each task type, appropriate prompts are generated and passed to the \gls{llm}. The responses are then processed using regex-based parsing methods specific to \gls{mcq} or numeric answers (after obtaining the relevant part of the response from the instruction-tuned models).
    The regex-based parsing is elaborate and also able to handle special cases such as scientific notation, or roman numerals.
    If the initial parsing is unsuccessful, the system employs an \gls{llm}-based extraction method as a fallback. The parsed or extracted answers then undergo metric computation and refusal detection.}
    \label{fig:process}
\end{figure}

\subsection{Performance as a function of molecular features}
To better understand if the performance of the models is correlated with specific features of the molecules, we analyzed the performance of the models as a function of the number of atoms and the complexity of the molecules.
\Cref{fig:correlation_plot_is_number_nmr_peaks_complexity} shows that the performance of the models is not correlated with the complexity of the molecules but rather with the number of atoms (\Cref{fig:correlation_plot_is_number_nmr_peaks_num_atoms}, similar trivial correlation for \Cref{fig:correlation_plot_is_electron_counts_num_atoms}).
The corresponding Spearman correlation coefficients are listed in \Cref{tab:correlation_coefficients}.

@@

\subsection{Influence of model scale}
\Cref{fig:model_size_plot} shows the performance of the models as a function of the number of parameters in the model.
We see that the performance of the models correlates with scale for the models of the LLama3 and Llama3.1 herd of models.

@@

\subsection{Human baseline}
\paragraph{App} To facilitate the collection of responses, we developed a responsive web application in Typescript using the Next.js\autocite{nextjs} app router framework.
This application handles serving the user interface and exposes various \gls{rest} \glspl{api} for relevant operations.
We utilize a Postgresql.
The web application is styled with Tailwind CSS\autocite{tailwindcss} using the shadcn/ui component library and uses NextAuth\autocite{nextauth} for easy and secure user authentication.
The application is hosted on the Vercel web hosting platform.

@@

We also recorded the time humans took to answer the questions. This time is the time from the question being displayed to the human to the human submitting the answer.
%Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (\Cref{fig:human_timing}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\)$\quad$, and \(p \approx \variable{output/spearman_experience_score_p.txt}\)).

@@

\subsection{Confidence estimates} \label{sec:confidence_estimates}

@@

\subsection{Leaderboard}
Our leaderboard is based on the tool chain developed for Matbench.\autocite{Dunn_2020}
Briefly, the \chembench pipeline produces standardized files in \texttt{json} format that contributors can add via pull requests to the \chembench repository.
The Markdown tables and interactive plots are automatically generated and updated on the \chembench website.

>>>>

\clearpage
\subsection{Benchmark corpus}
To ensure maximal interoperability with existing benchmarks or tools, we curated the data in an extended form of the widely used BigBench format.\autocite{srivastava2022beyond}
This also implies that future baselines can be built on top of our infrastructure if saved in the same format.

%\Cref{fig:flesch_kincaid_reading_ease} shows the distribution of the Flesch-Kincaid reading ease scores of the questions.
%We see that the questions are generally complex to read.

@@

        \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.}
    \scriptsize
    \begin{tabularx}{\textwidth}{p{2.7cm}p{4em}X}

@@

\normalsize

@@

\begin{table*}[!t]
    \centering
        \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.}
    \scriptsize
    \begin{tabularx}{\textwidth}{p{2.7cm}p{4em}X}
        \toprule
        Required Skills & Number of Questions & Example Question \\
        \midrule
        Knowledge & 770 & Which of the following salts is an acidic salt? \\
        & & A. \ce{NH4Cl} \\
        & & B. \ce{Na2CO3} \\
        & & C. \ce{NaH2PO4} \\
        & & D. \ce{Zn(OH)Cl} \\
        \midrule
        Reasoning & 825 & How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
        & & A. By measuring atomic force microscopy \\
        & & B. By measuring fourier transform infrared spectroscopy \\
        & & C. By measuring photoluminescence spectrum \\
        & & D. By measuring the absorption spectroscopy \\
        \midrule
        Intuition & 1001 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality). Which of the following two candidates would you prefer for further development? \\
        & & A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
        & & B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC(=O)c2ccco2)CC1[END\_SMILES] \\
        \midrule
        Calculation & 91 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution. The volume of \ce{KOH} solution required to reach the endpoint was \SI{6.81}{cm^3}. What is the average number of monomer units in each polymer chain of this sample? \\
        \bottomrule
    \end{tabularx}
    \label{tab:chembench_corpus_cognitive}
\end{table*}

\normalsize

@@

\begin{table}
    \caption{\textbf{Performance of the models on the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
    \resizebox{\textwidth}{!}{
    \variable{output/performance_table.tex}
    }
    \label{tab:performance_table}
\end{table}

\begin{table}
    \caption{\textbf{Performance of the models on the \enquote{tiny} subset of the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
    \resizebox{\textwidth}{!}{
    \variable{output/performance_table_human_subset.tex}
    }
    \label{tab:performance_table_human_subset}
\end{table}

@@

\subsection{Performance as a function of molecular features}
To better understand if the performance of the models is correlated with specific features of the molecules, we analyzed the performance of the models as a function of the number of atoms and the complexity of the molecules.
\Cref{fig:correlation_plot_is_number_nmr_peaks_complexity} shows that the performance of the models is not correlated with the complexity of the molecules but rather with the number of atoms (\Cref{fig:correlation_plot_is_number_nmr_peaks_num_atoms}, similar trivial correlation for \Cref{fig:correlation_plot_is_electron_counts_num_atoms}).
%The corresponding Spearman correlation coefficients are listed in \Cref{tab:correlation_coefficients}.

@@

\clearpage
\subsection{Influence of model scale}
\Cref{fig:model_size_plot} shows the performance of the models as a function of the number of parameters in the model.
We see that the performance of the models correlates with scale for the models of the LLama-3 and Llama-3.1 herd of models.

@@

\clearpage
\subsection{Implementation}
An overview of the benchmarking pipeline implemented in \chembench is shown in \cref{fig:process}. More detailed information can be found in the online documentation of the \chembench package at \url{https://lamalab-org.github.io/chem-bench/}.
\begin{figure}
    \centering
    \includegraphics[height=.8\textheight]{figures/process.pdf}
    \caption{\textbf{Overview of the benchmarking pipeline implemented in \chembench.} The process begins with JSON files containing task data, which are used to create \texttt{Task} objects and initialize the \texttt{ChemBenchmark}. \texttt{Prompter} objects are then created to handle different types of prompts for instruction-tuned and completion models.
    The \texttt{Task} Iterator differentiates between \gls{mcq} and other question types. For each task type, appropriate prompts are generated and passed to the \gls{llm}. The responses are then processed using regex-based parsing methods specific to \gls{mcq} or numeric answers (after obtaining the relevant part of the response from the instruction-tuned models).
    The regex-based parsing is elaborate and also able to handle special cases such as scientific notation, or roman numerals.
    If the initial parsing is unsuccessful, the system employs an \gls{llm}-based extraction method as a fallback. The parsed or extracted answers then undergo metric computation and refusal detection.}
    \label{fig:process}
\end{figure}

\clearpage
\subsection{Human baseline} \label{sec:human_baseline}
\paragraph{App} To facilitate the collection of responses, we developed a responsive web application in Typescript using the Next.js\autocite{nextjs} app router framework.
This application handles serving the user interface and exposes various \gls{rest} \glspl{api} for relevant operations.
We utilize a Postgresql.
The web application is styled with Tailwind CSS\autocite{tailwindcss} using the shadcn/ui component library and uses NextAuth\autocite{nextauth} for easy and secure user authentication.
The application is hosted on the Vercel web hosting platform.

@@

We also recorded the time humans took to answer the questions. This time is the time from the question being displayed to the human to the human submitting the answer.

@@

\clearpage
\subsection{Confidence estimates} \label{sec:confidence_estimates}

@@

\clearpage
\subsection{Leaderboard}
Our leaderboard is based on the tool chain developed for Matbench.\autocite{Dunn_2020}
Briefly, the \chembench pipeline produces standardized files in \texttt{json} format that contributors can add via pull requests to the \chembench repository.
The Markdown tables and interactive plots are automatically generated and updated on the \chembench website.

### 2024-10-18 src/tex/ms.tex
Update ms.tex
more cleanup

\begin{figure}[!htb]
    \centering
    \includegraphics[width=\textwidth]{figures/cb_corpus_v1.pdf}
    \caption{\textbf{Distribution of topics and required skills.} This circular plot illustrates the distribution of questions across various chemistry topics, along with the primary skills required to address them. The topics were manually classified, showing a varied representation across different aspects of chemistry. Each topic is associated with a combination of three key skills: Calculation, Reasoning, and Knowledge, as indicated by the colored bars. \chembench samples diverse topics and diverse skills, setting a high bar for \glspl{llm} to demonstrate human-competitive performance across a wide range of chemistry tasks.}
    \label{fig:corpus}
\end{figure}

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/human_subset_performance.pdf}
    \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered correctly. We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance. 
    Since the humans did not answer all the questions, this plot is based on the subset of questions that all humans answered.
    The evaluation we use here is very strict as it only considers a question answered correctly or incorrectly, partially correct answers are also considered incorrect.
    \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
    PaperQA2\autocite{skarlinski2024language} is an agentic system that can also search the literature to obtain an answer. We find that the best models outperform all humans in our study when averaged over all questions (even though humans had access to tools such as web search and ChemDraw for a subset of the questions). 
    }
    \label{fig:human_vs_models_bar}
    \script{plot_overview_performance_plot.py}
\end{figure}

% ToDO: Link to full table of results in the SI
\paragraph{System performance} 
To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\autocite{Huggingface} on the \chembench corpus, including systems augmented with external tools.
An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar}. 
In this figure, we show the percentage of questions that the models answered correctly.
Moreover, we show the worst, best, and average performance of the human experts in our study, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
Remarkably, the figure shows that the leading \gls{llm}, \oone, outperforms the best human in our study in this overall metric by almost a factor of two.
Many other models also outperform the average human performance. 
Interestingly, \LlamaThreeOneFourZeroFiveBInstruct shows performance that is close the leading proprietary models, indicating that new open source models can be competitive with the best proprietary models also in chemical settings.

Given the considerable interest in tool-augmented systems, the mediocre performance of these systems (GPT-3.5 and Claude 2 augmented with tools) in our benchmark is striking. 
Their lack of performance, however, is partially because we limited the system to a maximum of ten \gls{llm} calls.
With the default tool augmented setup (using the so-called ReAct method\autocite{yao2022react}), however, this often did not allow the system to identify the correct answer (e.g., because it repeatedly tried to search for the answer on the web and did not find a solution within ten calls to the \gls{llm}).
This observation highlights the importance of communicating and measuring not only predictive performance but also computational cost (e.g., in terms of \gls{api} calls) for tool-augmented systems.

% Importance of scale 
% Still lacks in knowledge-intensive questions

\paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}). We classified all questions in the \chembench corpus into these topics based on hand-crafted rules and classifier models operating on the question texts.
We then computed the percentage of questions the models or humans answered correctly for each topic.
\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_human.pdf}
    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus. The performance is measured as the fraction of questions that were answered completely correctly by the models.
    The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered correctly). A larger colored area indicates a better performance.
    This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
    }
    \label{fig:all_questions_models_completely_correct_radar_human}
    \script{analyze_model_reports.py}
\end{figure}

% In this spider chart, the worst score for every dimension is zero (no question answered correctly), and the best score is one (all questions answered correctly). 
% Thus, a larger colored area indicates a better performance. 
% One can observe that this performance varies widely across models and topics. 
% While macromolecular chemistry and biochemistry receive relatively high scores for many models, this is not the case for topics such as chemical safety or analytical chemistry.
% In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4) while this question appeared easier (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans.
% Importantly, the human experts are given a drawing of the compounds, whereas models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule). 
% These findings also shine an interesting light on the value of textbook-inspired questions. 
% A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students. 
% On those questions, the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
% For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt4.txt}\% correct answers for GPT-4, \variable{output/subset_scores/is_gfk_claude3.txt}\% for Claude 3, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
% While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning.

@@

REWRITE

%To investigate this, we prompted\autocite{xiong2023llms} some of the top-performing models to estimate, on an ordinal scale, their confidence in their ability to answer the question correctly.

@@

%\input{sections/semi_programatically_sources_table.tex}

\subparagraph{Chemical Preference data}

@@

\subsection{Confidence estimate}
To estimate the models' confidence, we prompted them with the question (and answer options for \gls{mcq}) and the task to rate their confidence to produce the correct answer on a scale from 1 to 5. 
We decided to use verbalized confidence estimates\autocite{xiong2023llms} since we found those closer to current practical use cases than other prompting strategies, which might be more suitable when implemented in systems.

@@

\paragraph{Study design}
Rewrite.

\paragraph{Participants}
Users were open to reporting about their experience in chemistry. 
Overall, \variable{output/num_users_with_education_info.txt} did so. 
Out of those, \variable{output/num_human_phd.txt} reported to have been awarded a Ph.D.
\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree, and \variable{output/num_human_bachelor.txt} have a bachelor's degree.

@@

\subsection{Data annotation}\label{sec:meth-topic} 
% When curating our dataset, we systematically recorded keywords and sources.
% To allow for analysis of the model performance as a function of the topic, we leverage this information together with the output of sequence classification models.
% We use this information to make the assignment for questions that can easily be assigned to a topic based on the source (e.g., number of \gls{nmr} signals, chemical compatibility, toxicology exam questions).
% For the remaining ones, e.g., from chemistry olympiad questions, we use zero-short sequence classification\autocite{zeroshotsequence} using the BART model\autocite{bart, FacebookBART}, which our preliminary analysis found to be more robust than topic modeling based on embeddings from OpenAI's \texttt{ada} model or Cohere's \texttt{Cohembed-english-v3.0} model.

REWRITE.

@@

@@

\scriptsize
\insertcredits
\normalsize
\printbibliography
\end{refsection}

>>>>

\begin{figure}[!htb]
    \centering
    \includegraphics[width=\textwidth]{figures/cb_corpus_v2.pdf}
    \caption{\textbf{Distribution of topics and required skills.} This circular plot illustrates the distribution of questions across various chemistry topics, along with the primary skills required to address them. The topics were manually classified, showing a varied representation across different aspects of chemistry. Each topic is associated with a combination of three key skills: Calculation, Reasoning, and Knowledge, as indicated by the colored bars. \chembench samples diverse topics and diverse skills, setting a high bar for \glspl{llm} to demonstrate human-competitive performance across a wide range of chemistry tasks.}
    \label{fig:corpus}
\end{figure}

@@

\begin{figure}[!h]
    \centering
    \includegraphics{figures/human_subset_performance.pdf}
    \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered correctly. We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance.
    Since the humans did not answer all the questions, this plot is based on the subset of questions that all humans answered.
    The evaluation we use here is very strict as it only considers a question answered correctly or incorrectly, partially correct answers are also considered incorrect.
    \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
    PaperQA2\autocite{skarlinski2024language} is an agentic system that can also search the literature to obtain an answer. We find that the best models outperform all humans in our study when averaged over all questions (even though humans had access to tools such as web search and ChemDraw for a subset of the questions).
    }
    \label{fig:human_vs_models_bar}
    \script{plot_overview_performance_plot.py}
\end{figure}

\paragraph{Overall system performance}
To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\autocite{Huggingface} on the \chembench corpus, including systems augmented with external tools.
An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar} (all results can be found in \Cref{tab:performance_table} and \Cref{tab:performance_table_human_subset}).
In this figure, we show the percentage of questions that the models answered correctly.
Moreover, we show the worst, best, and average performance of the human experts in our study, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
Remarkably, the figure shows that the leading \gls{llm}, \oone, outperforms the best human in our study in this overall metric by almost a factor of two.
Many other models also outperform the average human performance.
Interestingly, \LlamaThreeOneFourZeroFiveBInstruct shows performance that is close the leading proprietary models, indicating that new open source models can be competitive with the best proprietary models also in chemical settings.

Interestingly, we find that models are still limited in their ability to answer knowledge-intensive questions. Importantly, this is not a limitation that could be overcome by simple application of \gls{rag} systems such as \PaperQATwo. This is likely because the required knowledge cannot easily be accessed via papers but rather by lookup in specialized databases (e.g., PubChem, Gestis), which also the humans in our study used to answer such questions.
This indicates that there is still room for improvement of chemical \gls{llm} by training them on more specialized data sources or by integrating them with specialized databases.

In addition, our analysis shows that the performance of models is correlated with their size (see \Cref{fig:model_size_plot}). This is in line with observations in other domains but also indicates that chemical \gls{llm} could be further improved by scaling them up.

\paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}). We classified all questions in the \chembench corpus into these topics.
We then computed the percentage of questions the models or humans answered correctly for each topic and show them in \Cref{fig:all_questions_models_completely_correct_radar_human}.
In this spider chart, the worst score for every dimension is zero (no question answered correctly), and the best score is one (all questions answered correctly). Thus, a larger colored area indicates a better performance.

\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_human.pdf}
    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus. The performance is measured as the fraction of questions that were answered completely correctly by the models.
    The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered correctly). A larger colored area indicates a better performance.
    This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
    }
    \label{fig:all_questions_models_completely_correct_radar_human}
    \script{analyze_model_reports.py}
\end{figure}

One can observe that this performance varies across models and topics.
While general and technical chemistry and biochemistry receive relatively high scores for many models, this is not the case for topics such as toxicity and safety or analytical chemistry.

In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_o1.txt} percent correct answers for \oone) while this question appeared easier (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans.
 Importantly, the human experts are given a drawing of the compounds, whereas models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).
% These findings also shine an interesting light on the value of textbook-inspired questions.
% A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students.
% On those questions, the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
% For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt4.txt}\% correct answers for GPT-4, \variable{output/subset_scores/is_gfk_claude3.txt}\% for Claude 3, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
% While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning.

@@

To investigate this, we prompted\autocite{xiong2023llms} some of the top-performing models to estimate, on an ordinal scale, their confidence in their ability to answer the question correctly (see \Cref{sec:confidence_estimates} for details on methodology and comparison to logit-based approaches).

@@

\input{sections/semi_programatically_sources_table.tex}

\subparagraph{Chemical preference data}

@@

\subsection{Confidence estimate}
To estimate the models' confidence, we prompted them with the question (and answer options for \gls{mcq}) and the task to rate their confidence to produce the correct answer on a scale from 1 to 5.
We decided to use verbalized confidence estimates\autocite{xiong2023llms} since we found those closer to current practical use cases than other prompting strategies, which might be more suitable when implemented in systems. In addition, this approach captures semantic uncertainty which is not the same as the probability of a token given a sequence of tokens (i.e., the uncertainty one obtains from logit-based approaches).

@@

\paragraph{Study design}
Human participants were asked the questions in a custom-built web interface (\Cref{sec:human_baseline}) which rendered chemicals and equations. Questions were shown in random order and participants were not allowed to skip questions. For a subset of the questions, the participants were allowed to use external tools (excluding other \gls{llm} or asking other people) to answer the questions. Prior to answering questions, participants were asked to provide information about their education and experience in chemistry. The study was conducted in English.

\paragraph{Participants}
Users were open to reporting about their experience in chemistry.
Overall, \variable{output/num_users_with_education_info.txt} did so.
Out of those,
\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree (and are currently enrolled in Ph.D.\ studies), and \variable{output/num_human_bachelor.txt} have a bachelor's degree.

@@

\subsection{Data annotation}\label{sec:meth-topic}
In the curation of our dataset we manually assigned difficulty levels as well as required skills to each question. We used following guidelines for these annotations: calculation is required if answering a question would require the use of a calculator, knowledge is required if answering a question requires non-trivial knowledge of facts (e.g. the H/P statements of chemicals). Reasoning is required if answering a question requires multiple reasoning steps.
Basic questions are those that only requires those skills up to high-school level. Advanced questions would require an expert multiple minutes up to hours to answer.

@@

\section{Statement of ethical compliance}
The authors confirm to have complied with all relevant ethical regulations, according to the Ethics Commission of the Friedrich Schiller University Jena (which decided that study is ethically safe). Informed consent was obtained from all participants.

@@

\resizebox{\textwidth}{!}{%
\scriptsize
\insertcredits
}
\normalsize
\printbibliography
\end{refsection}
